---
title: "DSA2101 Group Project"
output: html_document
---
#### Group name: sixpeople

# Introduction
<p align="justify">
The UNVotes dataset compiled by tidytuesday comprises data about the voting history of 
various countries on UN resolutions and amendments since the first meeting in 1946. Votes have been categorised using six issues: Colonialism, Economic Development, Palestinian Conflict, Arms Control and Disarmament, Human Rights and Nuclear Weapons and Nuclear material. 

</p>
<p align="justify">
The key question that our group intends to explore with this dataset is: **How has the UN’s voting patterns evolved over time, and what notable trends and times of division can be observed? How closely do countries’ votes on these issues align with major powers such as the United States of America, China and Russia?** By exploring these trends, we aim to reveal how international alliances change and where consensus or division is present and how a country might change its stance across time. 
</p>

# Data Cleaning & Summary
<p align="justify">
We examined the **three separate datasets** in the UNVotes data: **un_votes** (countries’ voting records), **un_roll_calls** (information on each resolution) and **un_roll_call_issues** (showing the topic tags associated with each resolution). All three datasets shared a common key: **rcid**, which was then utilised in a series of left joins to merge the three datasets into one. This allowed us to match a resolution with its metadata (like date and description) and issue classifications. Additionally, we extracted the years from every resolution date using the mutate function in order to conduct analyses based on time. Below are the final columns used for our data analysis.
<p>

**rcid**: Unique roll call ID for each resolution  
**country**: Name of the country casting the vote  
**country_code**: ISO alpha-2 country code (e.g., “US”, “CA”)  
**vote**: Factor with 3 levels (“yes”, “no”, “abstain”)  
**session**: Numeric identifier for the UN session  
**importantvote**: Integer flag (0/1) for important votes  
**date**: Date the vote took place (format: “1946-01-01”)  
**amend**: Integer flag (0/1) for amendments  
**para**: Integer flag (0/1) for paragraph votes  
**short**: Brief label for the resolution (e.g., “AMENDMENTS, RULES OF PROCEDURE”)  
**descr**: Longer description of the resolution (e.g., “TO ADOPT A CUBAN AMENDMENT TO THE UK PROPOSAL”)

<p align="justify">
Our exploration of voting trends from 1946 to 2020 revealed several insights. In the earlier decades, Colonialism was the key issue, taking up a significant proportion of UN resolutions. This aligned with the move towards decolonisation efforts during that period. With Cold War tensions in the 1970s, more of the resolutions targeted nuclear weapons and their disarmament. In the 1980s, however, there was a marked increase in resolutions about the Palestinian conflict as the situation worsened in the Middle East. We also saw the rise and fall in the number of resolutions targeting issues like economic development, arms control and global inequality which could be tied to geopolitical events over the years. While human rights issues gained more of a spotlight from the 1970s, it also emerged to be the most divisive issue with only an average vote consensus percentage of 73%. This means that, on average, each resolution targeting human rights issues saw more than a quarter of the countries opposing the majority vote. This relatively lower level of consensus among voting countries made the issue an interesting candidate for further analysis.
<p>

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(forcats)
library(scales)
library(sf)
library(unvotes)
library(lubridate)
library(plotly)
library(countrycode)
library(gridExtra)
library(grid)
```

```{r,echo=TRUE}

# left join all by rcid
votes_full <- un_votes %>%
  left_join(un_roll_calls, by = "rcid") %>%
  left_join(un_roll_call_issues, by = "rcid") %>%
  mutate(year = year(date))

#str(votes_full)

```

# Plots and Data Visualisation

```{r, echo=TRUE, fig.show='hide', message=FALSE, warning=FALSE}
res_by_year_issue <- votes_full %>%
  filter(!is.na(issue), issue != "NA") %>%
  distinct(rcid, year, issue) %>%
  count(year, issue)

# filter for human rights
#res_by_year_issue <- res_by_year_issue %>%
#  filter(issue == "Human rights") %>%
#  mutate(issue = fct_reorder(issue, n))

graph1 <- ggplot(res_by_year_issue, aes(x = year, y = n, fill = issue)) +
  geom_line(alpha = 0.85, size = 0.2, color = "white") +
  scale_x_continuous(breaks = seq(1945, 2020, by = 10)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "UN Resolutions by Issue Over Time",
    x = "Year", y = "Resolutions", fill = "Issue"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")

```


```{r, echo=TRUE}
issue_totals <- votes_full %>%
  distinct(rcid, issue) %>%
  filter(!is.na(issue)) %>%
  count(issue, sort = TRUE) %>%
  mutate(pct = n / sum(n), label = paste0(round(pct * 100), "%"))

graph2 <- ggplot(issue_totals, aes(x = "", y = n, fill = issue)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4) +
  labs(title = "Distribution of UN Resolutions by Issue (1945 - Present)") +
  theme_void(base_size = 14) +
  theme(legend.position = "right")

```


```{r, echo=TRUE, fig.width=14, fig.height=7}
common_fill <- scale_fill_brewer(palette = "Dark2", name = "Issue")

# stackedbarchart
p1 <- ggplot(res_by_year_issue, aes(x = year, y = n, fill = issue)) +
  geom_area(alpha = 0.9, color = "white", size = 0.2) +
  common_fill +
  scale_x_continuous(breaks = seq(1945, 2020, by = 10)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "UN Resolutions by Issue Over Time",
    x = "Year", y = "Number of Resolutions"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    legend.position = "none"  # No legend here
  )



# # piechart (not used as faceted line graphs show the percentage each type issue holds out of all types of issues per year)
# p2 <- ggplot(issue_totals, aes(x = "", y = n, fill = issue)) +
#   geom_col(width = 1, color = "white") +
#   coord_polar(theta = "y") +
#   geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4) +
#   common_fill +
#   labs(title = "Overall Distribution (1945–Present)") +
#   theme_void(base_size = 14) +
#   theme(
#     plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
#     legend.position = "none"  # No legend here
#   )


#distribution per year
votes_full1=votes_full%>%
  filter(!is.na(issue))%>%
  select(rcid,year,issue,importantvote)%>%
  group_by(year)%>%
  mutate(issues_per_year=n())%>%
  ungroup()%>%
  group_by(year,issue)%>%
  mutate(percent_per_year=(n()/issues_per_year) *100)%>%
  ungroup()%>%
  distinct(year,issue,issues_per_year,percent_per_year)
  # filter(issue %in% c("Colonialism","Human rights","Economic development"))

# votes_full1=votes_full%>%
#   filter(!is.na(issue))%>%
#   select(rcid,year,issue,importantvote)%>%
#   group_by(year,issue)%>%
#   count()

# interested_issues=c("Colonialism","Human rights","Economic Development")

p3 <- ggplot(votes_full1,aes(x=year,y=percent_per_year,color=issue))+
  geom_line(alpha=0.5)+
  facet_wrap(~issue,scales="free_y")+
  scale_color_brewer(palette = "Dark2") + 
  labs(title = "Proportion of resolutions per year by issue type",
       y="Proportion/%",
       x=" ")+
  theme(
    legend.position = "none"
  )
# legend plot
legend_plot <- ggplot(res_by_year_issue, aes(x = year, y = n, fill = issue)) +
  geom_area() +
  common_fill +
  theme(legend.position = "bottom")
tmp <- ggplot_gtable(ggplot_build(legend_plot))
leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
legend <- tmp$grobs[[leg]]

# title
title <- textGrob("UN Resolutions: Trends and Overall Composition", 
                  gp = gpar(fontface = "bold", fontsize = 20))

# arranging everything neatly
final_plot <- grid.arrange(
  title,
  arrangeGrob(p1, p3, ncol = 1, heights = c(2, 2)),
  legend,
  heights = c(0.1, 0.8, 0.1),
  nrow = 3
)


```





<p align="justify">
**Fig 1:** The **stacked area chart** (Year on the x-axis, Resolution Count on the y-axis, colored by Issue Category) was chosen to show how the UN’s voting patterns have changed over time. It clearly illustrates shifts in focus across different issues, the rise or fall in resolution activity, and key periods of division or consensus. This helps identify historical turning points and changing global priorities. 
<p>
<p align="justify">
The **pie chart** (Issue categories as segments, sized by their share of total resolutions) provides an overall view of which issues have received the most attention. This helps us understand the broader themes that have shaped UN decision-making and offers insight into how countries have aligned their votes over time, especially in relation to major powers like the United States, China, and Russia. 
<p>

<p align="justify">
From the stacked area chart, we can see notable surges in the number of resolutions in the mid-1970s coinciding with decolonisation movements, the 1973 oil crisis and the then-intensifying Cold War. This period was then followed by a plateau in the number of resolutions following the dissolution of the Soviet Union which led to another peak in the late 1990s due to humanitarian intervention efforts in Yugoslavia and Rwanda. This analysis shows how global priorities shifted over time from issues like colonialism and arms control (during the Cold War) to other issues like human rights and economic development in the 1990s. Recent surges around 2015 can be attributed to multiple crises like civil unrest in Syria, migration challenges in Europe, climate change mitigation negotiations and the adoption of UN Sustainable Development Goals. From the pie chart, we can also see that since its inception in 1946, arms control and the Palestinian conflict has remained dominant around evolving international priorities. 
<p>

```{r, echo=TRUE, fig.width=12, fig.height=6}

group_definitions <- tribble(
  ~country,               ~group,
  # G7
  "United States",        "G7",
  "United Kingdom",       "G7",
  "France",               "G7",
  "Germany",              "G7",
  "Italy",                "G7",
  "Canada",               "G7",
  "Japan",                "G7",
  
  # BRICS
  "Brazil",               "BRICS",
  "Russia",               "BRICS",
  "India",                "BRICS",
  "China",                "BRICS",
  "South Africa",         "BRICS",
  
  # NATO (subset to avoid G7 overlap)
  "Turkey",               "NATO",
  "Spain",                "NATO",
  "Portugal",             "NATO",
  "Greece",               "NATO",
  "Netherlands",          "NATO",
  "Belgium",              "NATO",
  "Poland",               "NATO",
  "Czech Republic",       "NATO",
  "Hungary",              "NATO",
  "Norway",               "NATO",
  "Denmark",              "NATO",
  "Iceland",              "NATO",
  "Luxembourg",           "NATO",
  
  
  # OPEC
  "Saudi Arabia",         "OPEC",
  "Iran",                 "OPEC",
  "Iraq",                 "OPEC",
  "Kuwait",               "OPEC",
  "Venezuela",            "OPEC",
  "United Arab Emirates", "OPEC",
  "Nigeria",              "OPEC",
  
  # ASEAN
  "Indonesia",            "ASEAN",
  "Thailand",             "ASEAN",
  "Malaysia",             "ASEAN",
  "Philippines",          "ASEAN",
  "Singapore",            "ASEAN",
  "Vietnam",              "ASEAN",
  "Myanmar",              "ASEAN",
  "Cambodia",             "ASEAN",
  "Laos",                 "ASEAN",
  "Brunei",               "ASEAN"
)
votes_hr <- votes_full %>%
  filter(issue == "Human rights") %>%
  select(rcid, year, country, vote) %>%
  left_join(group_definitions, by = "country")

us_votes <- votes_hr %>%
  filter(country == "United States") %>%
  select(rcid, us_vote = vote)

votes_compared <- votes_hr %>%
  filter(country != "United States") %>%
  left_join(us_votes, by = "rcid") %>%
  filter(!is.na(us_vote), !is.na(vote)) %>%
  mutate(aligned = (vote == us_vote))

group_alignment_over_time <- votes_compared %>%
  filter(!is.na(group)) %>%
  group_by(group, year) %>%
  summarise(
    avg_alignment = mean(aligned),
    n_votes = n(),
    .groups = "drop"
  )

ggplot(group_alignment_over_time, aes(x = year, y = avg_alignment, color = group)) +
  geom_point(size = 2, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, size = 1.2, span = 0.5) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Bloc Alignment with USA on Human Rights Votes",
    subtitle = "Smoothed trend of bloc agreement with US votes (Human Rights, 1945–present)",
    x = "Year",
    y = "Alignment with USA",
    color = "Bloc"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 13)
  )

```

```{r,echo=TRUE}
votes_us_yearly <- votes_full %>%
  filter(issue == "Human rights") %>% 
  filter(year >= 2005, year <= 2015) %>%
  select(rcid, country, vote, date, year) %>%
  left_join(
    un_votes %>% filter(country == "United States") %>% select(rcid, us_vote = vote),
    by = "rcid"
  ) %>%
  filter(!is.na(us_vote)) %>%
  mutate(
    agree_with_us = vote == us_vote,
    code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) %>%
  group_by(code, year) %>%
  summarise(
    alignment_score = mean(agree_with_us, na.rm = TRUE),
    .groups = "drop"
  )

```
<p align="justify">
**Fig 2:** We also conducted a time-series analysis (using a line chart) of international bloc alignments on human rights issues, using the US vote as a reference. For each resolution, we measured if a bloc’s vote aligned with the US (agreement = 1, disagreement = 0) and averaged these results by year and bloc. This allowed us to track the evolution of political alignment over time, addressing the main question of how countries' voting patterns align with major powers like the US.
<p>

<p align="justify">
The analysis from 1946 onwards shows similarities and deviations in trends among geopolitical groups. For instance, G7 and NATO members like Germany, France, and Japan tend to hold similar positions on human rights issues. In contrast, emerging geopolitical groups such as BRICS (Brazil, Russia, India, China, South Africa) and certain OPEC countries like Saudi Arabia and Iran generally show lower and more inconsistent alignment with Western nations. This indicates a shift in international views and priorities, where human rights issues may be overshadowed by concerns over sovereignty and national development. 
<p>


```{r,echo=TRUE, fig.width=10, fig.height=5}
p <- plot_geo(votes_us_yearly, locationmode = 'ISO-3') %>%
  add_trace(
    z = ~alignment_score,
    locations = ~code,
    frame = ~year,
    color = ~alignment_score,
    colors = "RdYlGn",
    zmin = 0,
    zmax = 1
  ) %>%
  colorbar(title = "Alignment with U.S.") %>%
  layout(
    title = "<b>Yearly Alignment with U.S. on UN Votes (2005–2015)</b>",
    geo = list(
      showframe = FALSE,
      showcoastlines = TRUE,
      projection = list(type = 'equirectangular')
    ),
    updatemenus = list(
      list(
        type = "buttons",
        showactive = FALSE,
        buttons = list(
          list(
            method = "animate",
            args = list(NULL, list(
              frame = list(duration = 1000, redraw = TRUE),
              fromcurrent = TRUE,
              transition = list(duration = 300)
            )),
            label = "Play"
          ),
          list(
            method = "animate",
            args = list(NULL, list(
              mode = "immediate",
              frame = list(duration = 0, redraw = FALSE),
              transition = list(duration = 0)
            )),
            label = "Pause"
          )
        )
      )
    )
  )


p
```


<p align="justify">
**Fig 3:** The interactive choropleth map visualizes each nation's alignment with the US on resolutions from 2005 to 2015, a period we narrowed down due to an unusually high number of resolutions passed during those years. The alignment score, calculated in the same way as for blocs but applied to individual countries, highlights evolving geopolitical alliances and shifts in voting patterns in response to global priorities.
<p>

<p align="justify">
It’s easy to see that countries within the same geographic regions often voted in similar ways, with noticeable patterns emerging across the identified geopolitical blocs. Although most Western countries generally align with the US on human rights issues, there was a sharp drop in agreement in 2008. Further analysis revealed that this was largely due to a surge in human rights resolutions related to the Israel-Palestine conflict during that year. This issue proved especially divisive even among Western blocs. Toward the end of the timeline in 2015, however, we observed a trend of increasing alignment with the US on human rights resolutions across more countries.
<p>

# Final Conclusion/Discussion
<p align="justify">
In conclusion, our analysis of UN voting trends from 1946 to 2020 reveals how global priorities and alliances evolved over time. Colonialism and nuclear disarmament were central early on, followed by an increased focus on the Palestinian conflict and human rights in the 1980s and 1990s. Human rights remained the most divisive issue, with only 73% average consensus. Geopolitical events like the fall of the Soviet Union and humanitarian crises influenced UN resolutions. Different blocs, such as G7, NATO, BRICS, and OPEC, showed varying alignments with the US, especially on human rights, with notable divisions like the 2008 Israel-Palestine conflict. These trends highlight the shifting dynamics of global cooperation and division.
<p>








